---
title: "IRC Latin America — Health Services Dashboard"
---

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(patchwork)
library(tidytext)

source("scripts/theme_irc.R")

last_updated_raw <- readLines("data/last_updated.txt")
date_str         <- str_extract(last_updated_raw, "[0-9]{4}-[0-9]{2}-[0-9]{2}")
last_updated_fmt <- paste0(
  "Last updated: ", format(as.Date(date_str), "%d %B %Y"),
  ". Source: CommCare, IRC Latin America."
)

# ── Data loading ───────────────────────────────────────────────────────────────

monthly        <- read_csv("data/agg_monthly_consultations.csv",    show_col_types = FALSE)
sex_bdown      <- read_csv("data/agg_sex_breakdown.csv",            show_col_types = FALSE)
age_bdown      <- read_csv("data/agg_age_breakdown.csv",            show_col_types = FALSE)
pyramid        <- read_csv("data/agg_age_sex_pyramid.csv",          show_col_types = FALSE)
migration      <- read_csv("data/agg_migration_profile.csv",        show_col_types = FALSE)
ethnicity      <- read_csv("data/agg_ethnic_profile.csv",           show_col_types = FALSE)
loc_counts     <- read_csv("data/agg_location_consultations.csv",   show_col_types = FALSE)
loc_coords     <- read_csv("data/locations.csv",                    show_col_types = FALSE)
aps_dx         <- read_csv("data/agg_aps_diagnoses.csv",            show_col_types = FALSE)
aps_age        <- read_csv("data/agg_aps_by_age.csv",               show_col_types = FALSE)
ssr_mix        <- read_csv("data/agg_ssr_service_mix.csv",          show_col_types = FALSE)
its_results    <- read_csv("data/agg_its_results.csv",              show_col_types = FALSE)
prenatal_age   <- read_csv("data/agg_prenatal_age.csv",             show_col_types = FALSE)
cacu           <- read_csv("data/agg_cacu_screening.csv",           show_col_types = FALSE)
mh_service     <- read_csv("data/agg_mh_service_type.csv",          show_col_types = FALSE)
mh_dx          <- read_csv("data/agg_mh_diagnoses.csv",             show_col_types = FALSE)
referrals      <- read_csv("data/agg_referrals.csv",                show_col_types = FALSE)
equity_dx      <- read_csv("data/agg_equity_diagnoses.csv",         show_col_types = FALSE)
its_by_pop     <- read_csv("data/agg_its_by_population.csv",        show_col_types = FALSE)
prenatal_pop   <- read_csv("data/agg_prenatal_by_population.csv",   show_col_types = FALSE)
migration_time <- read_csv("data/agg_migration_by_month.csv",       show_col_types = FALSE)
its_time       <- read_csv("data/agg_its_by_month.csv",             show_col_types = FALSE)
aps_trend      <- read_csv("data/agg_aps_trend.csv",                show_col_types = FALSE)
prenatal_trend <- read_csv("data/agg_prenatal_trend.csv",           show_col_types = FALSE)
mh_time        <- read_csv("data/agg_mh_by_month.csv",              show_col_types = FALSE)

label_country <- function(x) str_to_title(x)
cutoff        <- as.Date("2024-05-01")

# ── ICD-10 label translations ──────────────────────────────────────────────────
dx_lookup <- c(
  B829 = "Intestinal parasitosis",    N390 = "Urinary tract infection",
  J00  = "Common cold",               Z001 = "Child health check",
  R51  = "Headache",                  J039 = "Acute tonsillitis",
  Z718 = "Other specified consultation", J22 = "Acute lower resp. infection",
  B86  = "Scabies",                   M545 = "Low back pain",
  Z308 = "Contraception care",        J029 = "Acute pharyngitis",
  K029 = "Dental caries",             K297 = "Gastritis",
  J40  = "Bronchitis",                D509 = "Iron deficiency anaemia",
  J459 = "Asthma",                    F99  = "Mental disorder, unspecified",
  Z659 = "Psychosocial problems",     Z639 = "Primary support group problem",
  F419 = "Anxiety disorder, unspecified", F320 = "Mild depressive episode",
  F439 = "Acute stress reaction",     F329 = "Depressive episode, unspecified",
  F321 = "Moderate depressive episode", Z609 = "Social environment problem",
  F322 = "Severe depressive episode", F431 = "PTSD",
  Z619 = "Adverse experience problem", R448 = "Perceptual disturbance",
  Z729 = "Lifestyle problem",         Z625 = "Child neglect problem",
  R453 = "Demoralisation / apathy",   Z733 = "Stress-related problem"
)

label_dx <- function(code, label, maxlen = 32) {
  translated <- dx_lookup[code]
  coalesce(translated, str_trunc(str_to_sentence(label), maxlen))
}

# ── Shared label lookups ──────────────────────────────────────────────────────
condition_labels <- c(sifilis = "Syphilis", vih = "HIV", hepatitis = "Hepatitis B")

pop_labels <- c(local_host = "Host community", migrant_refugee = "Migrants / refugees")

migration_labels <- c(
  poblacion_acogida                     = "Host community",
  migrante_en_transito                  = "Transit migrant",
  migrante_asentado                     = "Settled migrant",
  "migrante_con_vocacin_de_permanencia" = "Migrant (permanent intent)",
  local_transito                        = "Local (in transit)",
  refugiado                             = "Refugee / asylum seeker",
  desplazado_interno                    = "Internally displaced",
  retornado                             = "Returnee / deportee"
)

# ── Precomputed headline stats ─────────────────────────────────────────────────
pct_under18 <- age_bdown %>%
  filter(age_band %in% c("0-5", "6-11", "12-17")) %>%
  summarise(pct = sum(n) / sum(age_bdown$n)) %>% pull(pct)

pct_indigena <- ethnicity %>%
  filter(ethnic_profile == "indigena") %>%
  summarise(pct = sum(n) / sum(ethnicity$n)) %>% pull(pct)

n_total <- monthly %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff) %>%
  pull(n_consultations) %>% sum()

pct_prenatal_under18 <- prenatal_age %>%
  mutate(is_u18 = age_band %in% c("0-5", "6-11", "12-17")) %>%
  summarise(pct = sum(n[is_u18]) / sum(n)) %>% pull(pct)

syphilis_combined_positivity <- its_results %>%
  filter(condition == "sifilis") %>%
  summarise(rate = sum(n_positive) / sum(n_tested)) %>% pull(rate)
```

`r last_updated_fmt`

---

## Section 0: Geographic overview

#### Chart 0.1 — Consultation locations

Where IRC health services are being provided across Ecuador and Peru. Bubble size is proportional to total registered clients at each municipality.

```{r chart-0-1}
#| fig-width: 9
#| fig-height: 7

map_data <- loc_counts %>%
  left_join(select(loc_coords, municipio, lat, lon), by = "municipio") %>%
  filter(!is.na(lat), !is.na(lon)) %>%
  mutate(country_label = label_country(country))

region_sf <- ne_countries(
  continent = c("South America", "North America"),
  scale = "medium", returnclass = "sf"
)
locs_sf <- st_as_sf(map_data, coords = c("lon", "lat"), crs = 4326)

ggplot() +
  geom_sf(data = region_sf, fill = "grey94", colour = "grey72", linewidth = 0.2) +
  geom_sf(data = locs_sf, aes(size = n_clients, colour = country_label), alpha = 0.8) +
  scale_size_area(max_size = 16, guide = "none") +
  scale_colour_country(name = NULL) +
  coord_sf(xlim = c(-83, -66.5), ylim = c(-14.5, 3.0), expand = FALSE) +
  labs(caption = "Bubble size = total registered clients. Municipality centroids, manually verified.") +
  theme_irc() +
  theme(
    axis.text = element_blank(), axis.ticks = element_blank(),
    axis.title = element_blank(), panel.grid = element_blank(),
    legend.position = "bottom"
  )
```

---

## Section 1: Who we're reaching

#### Chart 1.1 — Consultations over time

```{r chart-1-1}
#| fig-width: 9
#| fig-height: 4

monthly_clean <- monthly %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff) %>%
  mutate(country = label_country(country))

ggplot(monthly_clean, aes(x = date, y = n_consultations, colour = country)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(size = 1.8, alpha = 0.8) +
  scale_colour_country() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  labs(
    x = NULL, y = "Consultations", colour = NULL,
    caption = paste0(
      "N = ", comma(n_total), " consultations (May 2024 onwards). ",
      "Earlier records excluded as likely test data."
    )
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Chart 1.2 — Sex breakdown

```{r chart-1-2}
#| fig-width: 9
#| fig-height: 3.5

sex_labels <- c(H = "Men", M = "Women", I = "Intersex", NR = "Prefer not to say")

sex_plot <- sex_bdown %>%
  mutate(
    sex_label = sex_labels[sex],
    sex_label = factor(sex_label, levels = c("Women", "Men", "Intersex", "Prefer not to say")),
    country   = label_country(country)
  )

ggplot(sex_plot, aes(x = pct, y = sex_label)) +
  geom_col(fill = "grey60", alpha = 0.8) +
  facet_wrap(~ country, ncol = 3) +
  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    x = NULL, y = NULL,
    caption = "Sex recorded at registration, not gender."
  )
```

#### Chart 1.3 — Age and sex distribution

`r percent(pct_under18, accuracy = 1)` of all registered clients are under 18.

```{r chart-1-3}
#| fig-width: 9
#| fig-height: 5.5

age_bands_5yr  <- c("0-4","5-9","10-14","15-19","20-24","25-29",
                    "30-34","35-39","40-44","45-49","50-54","55-59","60-64","65+")
age_labels_5yr <- str_replace(age_bands_5yr, "-", "–")
names(age_labels_5yr) <- age_bands_5yr

pyramid_plot <- pyramid %>%
  mutate(
    sex_label = if_else(sexo == "M", "Women", "Men"),
    value     = if_else(sexo == "M", pct, -pct),
    age_5yr   = factor(age_5yr, levels = age_bands_5yr),
    country   = label_country(country)
  )
x_limit <- max(abs(pyramid_plot$value)) * 1.08

ggplot(pyramid_plot, aes(x = value, y = age_5yr, fill = sex_label)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2) +
  scale_x_continuous(limits = c(-x_limit, x_limit),
                     labels = function(x) percent(abs(x), accuracy = 0.1)) +
  scale_y_discrete(labels = age_labels_5yr) +
  scale_fill_sex() +
  labs(x = NULL, y = NULL, fill = NULL,
       caption = "% of total male/female clients. Intersex and prefer-not-to-say excluded (< 0.1% combined).") +
  theme(legend.position = "top")
```

#### Chart 1.4 — Migration profile

Ecuador's clients are predominantly host community (77%); Peru's are predominantly transit migrants (69%).

```{r chart-1-4}
#| fig-width: 9
#| fig-height: 4.5

migration_plot <- migration %>%
  mutate(
    profile_label = coalesce(migration_labels[migration_profile], migration_profile),
    country       = label_country(country)
  ) %>%
  group_by(country) %>%
  mutate(profile_label = reorder(profile_label, pct)) %>%
  ungroup()

ggplot(migration_plot, aes(x = pct, y = profile_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 3, scales = "free_y") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL)
```

#### Chart 1.5 — Ethnic profile

`r percent(pct_indigena, accuracy = 0.1)` of clients across Ecuador and Peru self-identify as indigenous. Not collected in the Mexico app.

```{r chart-1-5}
#| fig-width: 7
#| fig-height: 4

ethnic_labels <- c(
  mestizo                 = "Mestizo",
  indigena                = "Indigenous",
  afrodescendiente        = "Afro-descendant",
  caucsico                = "White / Caucasian",
  sin_autorreconocimiento = "No self-identification",
  montubio                = "Montubio",
  otro_ee                 = "Other",
  otro_ep                 = "Other"
)

ethnic_plot <- ethnicity %>%
  mutate(
    ethnic_label = coalesce(ethnic_labels[ethnic_profile], ethnic_profile),
    country      = label_country(country)
  ) %>%
  group_by(country) %>%
  mutate(ethnic_label = reorder(ethnic_label, pct)) %>%
  ungroup()

ggplot(ethnic_plot, aes(x = pct, y = ethnic_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2, scales = "free_y") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL)
```

---

## Section 2: Clinical picture — primary care

#### Chart 2.1 — Top 10 diagnoses

```{r chart-2-1}
#| fig-width: 10
#| fig-height: 6

dx_top10 <- aps_dx %>%
  mutate(dx_label = label_dx(diagnosis_code, diagnosis_label)) %>%
  group_by(country) %>%
  slice_max(n, n = 10) %>%
  mutate(
    dx_label = reorder(dx_label, n),
    country  = label_country(country)
  ) %>%
  ungroup()

ggplot(dx_top10, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = comma) +
  labs(
    x = "Consultations", y = NULL,
    caption = "Top 10 diagnoses per country by total consultation count. All time periods."
  )
```

#### Chart 2.2 — Top diagnoses by age band

```{r chart-2-2}
#| fig-width: 12
#| fig-height: 15

age_order <- c("0-5", "6-11", "12-17", "18-49", "50+")

dx_by_age <- aps_age %>%
  mutate(
    dx_label = label_dx(diagnosis_code, diagnosis_label),
    age_band = factor(age_band, levels = age_order),
    country  = label_country(country)
  ) %>%
  group_by(country, age_band) %>%
  slice_max(n, n = 5) %>%
  mutate(dx_label = reorder_within(dx_label, n, list(country, age_band))) %>%
  ungroup()

ggplot(dx_by_age, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_grid(age_band ~ country, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  labs(
    x = "Consultations", y = NULL,
    caption = "Top 5 diagnoses per country and age band."
  ) +
  theme(
    axis.text.y  = element_text(size = 8),
    strip.text.y = element_text(size = 9)
  )
```

#### Chart 2.3 — Chronic disease prevalence

*Data not yet available.* The `patologias_cronicas` field was stripped by an overly broad anonymisation pattern. The fix is in place; re-running `00_anonymise_epr.R` will recover this field before the next pipeline run.

---

## Section 3: Sexual and reproductive health

#### Chart 3.1 — SSR service mix

```{r chart-3-1}
#| fig-width: 8
#| fig-height: 4.5

ssr_labels <- c(
  planificacin_familiar = "Family planning",
  atencin_prenatal      = "Prenatal care",
  its                   = "ITS/STI",
  despistaje_de_cancer  = "Cancer screening",
  vbg                   = "GBV services",
  consejera_en_ssr      = "SSR counselling"
)

ssr_plot <- ssr_mix %>%
  mutate(
    service_label = coalesce(ssr_labels[ssr_tipo], ssr_tipo),
    country       = label_country(country)
  )

ggplot(ssr_plot, aes(x = country, y = pct, fill = service_label)) +
  geom_col(alpha = 0.8) +
  scale_fill_manual(values = pal_service, name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = NULL,
    caption = "Proportion of SSR consultations by service type."
  ) +
  theme(legend.position = "right")
```

#### Chart 3.2 — ITS/STI testing and positivity

Syphilis positivity across both countries combined: **`r percent(syphilis_combined_positivity, accuracy = 0.1)`**. Peru (5.9%) is substantially higher than Ecuador (1.8%).

```{r chart-3-2}
#| fig-width: 8
#| fig-height: 4

its_plot <- its_results %>%
  mutate(
    condition_label = coalesce(condition_labels[condition], condition),
    condition_label = factor(condition_label, levels = c("Syphilis", "HIV", "Hepatitis B")),
    country         = label_country(country)
  )

ggplot(its_plot, aes(x = condition_label, y = positivity_rate, fill = country)) +
  geom_col(alpha = 0.8, position = position_dodge(width = 0.7), width = 0.6) +
  geom_text(
    aes(label = paste0("n=", n_tested), y = positivity_rate + 0.003),
    position = position_dodge(width = 0.7),
    size = 3, vjust = 0, family = "Inter"
  ) +
  scale_fill_country(name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1), limits = c(0, NA)) +
  labs(
    x = NULL, y = "Positivity rate",
    caption = "n = number tested. Hepatitis B sample sizes are small — interpret with caution."
  )
```

#### Chart 3.3 — Family planning method mix

*Data not available.* The `fp_corta` (short-acting) and `fp_larga` (long-acting) fields are empty across the full EPR dataset. Family planning consultations are counted in the SSR service mix above but method breakdown is not available pending investigation of the CommCare field names.

#### Chart 3.4 — Prenatal care: age profile

`r percent(pct_prenatal_under18, accuracy = 1)` of prenatal clients are under 18. Under-18 bars are highlighted in red.

```{r chart-3-4}
#| fig-width: 7
#| fig-height: 4

age_order <- c("0-5", "6-11", "12-17", "18-49", "50+")
age_display <- c("0–5", "6–11", "12–17", "18–49", "50+")
names(age_display) <- age_order

prenatal_plot <- prenatal_age %>%
  mutate(
    age_band  = factor(age_band, levels = age_order),
    bar_fill  = if_else(age_band %in% c("0-5", "6-11", "12-17"), col_highlight,
                        pal_country[str_to_title(country)]),
    country   = label_country(country)
  )

ggplot(prenatal_plot, aes(x = age_band, y = pct, fill = bar_fill)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_identity() +
  scale_x_discrete(labels = age_display) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    x = NULL, y = NULL,
    caption = "Red bars = under 18. Age at registration."
  )
```

#### Chart 3.5 — Cervical cancer screening (CACU/IVAA)

```{r chart-3-5}
#| output: asis

cacu_ecu <- cacu %>% filter(country == "ecuador")

cat(paste0(
  "**Ecuador:** ", cacu_ecu$n_screened_ivaa, " clients screened by IVAA. ",
  "Positivity rate: ", percent(cacu_ecu$ivaa_positivity_rate, accuracy = 0.1),
  " (", cacu_ecu$n_ivaa_positive, " positive).",
  "\n\n",
  "**Peru:** No IVAA screening records in current dataset. ",
  "Confirm whether the service runs in Peru or whether the field name differs between country apps."
))
```

---

## Section 4: Mental health

*Mental health services are provided in Peru and Ecuador. This module is not currently part of the Mexico programme.*

#### Chart 4.1 — Mental health service type

```{r chart-4-1}
#| fig-width: 8
#| fig-height: 3.5

mh_service_labels <- c(
  atencin_psicsocial_basada_en_la_comunidad                  = "Community psychosocial",
  atencin_psicosocial_focalizada                             = "Focused psychosocial",
  "atencin_especializada_referencia_a_atencin_especializada" = "Specialised / referred"
)

mh_svc_plot <- mh_service %>%
  mutate(
    svc_label = coalesce(mh_service_labels[service_type], service_type),
    svc_label = reorder(svc_label, pct),
    country   = label_country(country)
  )

ggplot(mh_svc_plot, aes(x = pct, y = svc_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL)
```

#### Chart 4.2 — Top mental health diagnoses

```{r chart-4-2}
#| fig-width: 9
#| fig-height: 5

mh_top <- mh_dx %>%
  mutate(dx_label = label_dx(diagnosis_code, diagnosis_label)) %>%
  group_by(country) %>%
  slice_max(n, n = 10) %>%
  mutate(
    dx_label = reorder(dx_label, n),
    country  = label_country(country)
  ) %>%
  ungroup()

ggplot(mh_top, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_wrap(~ country, ncol = 2, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = comma) +
  labs(x = "Consultations", y = NULL)
```

#### Chart 4.3 — External referral rate

```{r chart-4-3}
#| fig-width: 5
#| fig-height: 3

ref_plot <- referrals %>%
  mutate(country = label_country(country))

ggplot(ref_plot, aes(x = country, y = referral_rate, fill = country)) +
  geom_col(alpha = 0.8, width = 0.5) +
  geom_text(
    aes(label = paste0(percent(referral_rate, accuracy = 0.1),
                       "\n(", n_referred, " of ", comma(n_consultations), ")")),
    vjust = -0.3, size = 3.2, family = "Inter"
  ) +
  scale_fill_country(guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1),
                     limits = c(0, max(ref_plot$referral_rate) * 1.35)) +
  labs(x = NULL, y = "Referral rate",
       caption = "Proportion of mental health consultations resulting in external specialist referral.")
```

---

## Section 5: Pharmaceutical dispensing

*Pharmaceutical dispensing data is recorded in Mexico. The Mexico pipeline has not yet been built — this section will be populated in a future update.*

---

## Section 6: Equity lens

#### Chart 6.1 — Top diagnoses: host community vs migrants

```{r chart-6-1}
#| fig-width: 11
#| fig-height: 9

eq_plot <- equity_dx %>%
  mutate(
    dx_label       = label_dx(diagnosis_code, diagnosis_label),
    pop_label      = coalesce(pop_labels[population_group], population_group),
    country        = label_country(country)
  ) %>%
  group_by(country, pop_label) %>%
  mutate(dx_label = reorder_within(dx_label, n, list(country, pop_label))) %>%
  ungroup()

ggplot(eq_plot, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.8) +
  facet_grid(pop_label ~ country, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  labs(
    x = "Consultations", y = NULL,
    caption = "Top 5 diagnoses per country and population group."
  ) +
  theme(axis.text.y = element_text(size = 8))
```

#### Chart 6.2 — ITS positivity by population group

```{r chart-6-2}
#| fig-width: 9
#| fig-height: 4.5

its_pop_plot <- its_by_pop %>%
  mutate(
    condition_label = coalesce(condition_labels[condition], condition),
    condition_label = factor(condition_label, levels = c("Syphilis", "HIV", "Hepatitis B")),
    pop_label       = coalesce(pop_labels[population_group], population_group),
    country         = label_country(country)
  )

ggplot(its_pop_plot,
       aes(x = pop_label, y = positivity_rate, fill = country)) +
  geom_col(alpha = 0.8, position = position_dodge(0.7), width = 0.6) +
  geom_text(
    aes(label = paste0("n=", n_tested), y = positivity_rate + 0.005),
    position = position_dodge(0.7), size = 2.8, vjust = 0, family = "Inter"
  ) +
  facet_grid(condition_label ~ country, scales = "free_y") +
  scale_fill_country(guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(
    x = NULL, y = "Positivity rate",
    caption = "n = number tested. Many cells have small sample sizes — interpret with caution."
  ) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

#### Chart 6.3 — Proportion of prenatal clients under 18 by population group

```{r chart-6-3}
#| fig-width: 7
#| fig-height: 3.5

prenatal_pop_plot <- prenatal_pop %>%
  mutate(
    pop_label = population_group,
    country   = label_country(country)
  )

ggplot(prenatal_pop_plot,
       aes(x = pop_label, y = pct_under_18, fill = country)) +
  geom_col(alpha = 0.8, width = 0.5) +
  geom_text(
    aes(label = paste0(percent(pct_under_18, accuracy = 1), "\n(n=", n_prenatal, ")")),
    vjust = -0.3, size = 3, family = "Inter"
  ) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_country(guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1),
                     limits = c(0, max(prenatal_pop_plot$pct_under_18) * 1.4)) +
  labs(
    x = NULL, y = "% under 18",
    caption = "Age at registration."
  )
```

---

## Section 7: Trends over time

#### Chart 7.1 — Migration profile over time

```{r chart-7-1}
#| fig-width: 11
#| fig-height: 6

top_profiles <- migration_time %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff) %>%
  group_by(migration_profile) %>%
  summarise(total = sum(n)) %>%
  slice_max(total, n = 6) %>%
  pull(migration_profile)

mig_trend <- migration_time %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff, migration_profile %in% top_profiles) %>%
  mutate(
    profile_label = coalesce(migration_labels[migration_profile], migration_profile),
    country       = label_country(country)
  )

ggplot(mig_trend, aes(x = date, y = n, colour = country)) +
  geom_point(size = 2, alpha = 0.8) +
  facet_wrap(~ profile_label, ncol = 3, scales = "free_y") +
  scale_colour_country() +
  scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +
  scale_y_continuous(labels = comma) +
  labs(
    x = NULL, y = "Clients", colour = NULL,
    caption = "Top 6 migration categories by total volume. May 2024 onwards."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Chart 7.2 — ITS positivity over time

```{r chart-7-2}
#| fig-width: 9
#| fig-height: 4

its_trend <- its_time %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff) %>%
  mutate(
    condition_label = coalesce(condition_labels[condition], condition),
    condition_label = factor(condition_label, levels = c("Syphilis", "HIV", "Hepatitis B")),
    country         = label_country(country)
  ) %>%
  filter(n_tested >= 5)

ggplot(its_trend, aes(x = date, y = positivity_rate, colour = country)) +
  geom_point(aes(size = n_tested), alpha = 0.8) +
  facet_wrap(~ condition_label, ncol = 3) +
  scale_colour_country() +
  scale_size_area(max_size = 6, guide = "none") +
  scale_x_date(date_labels = "%b %y", date_breaks = "2 months") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1), limits = c(0, NA)) +
  labs(
    x = NULL, y = "Positivity rate", colour = NULL,
    caption = "Point size = number tested. Months with < 5 tests excluded."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Chart 7.3 — Top APS diagnosis trends

```{r chart-7-3}
#| fig-width: 10
#| fig-height: 6

aps_trend_plot <- aps_trend %>%
  mutate(
    date     = ym(year_month),
    dx_label = label_dx(diagnosis_code, diagnosis_label),
    country  = label_country(country)
  ) %>%
  filter(date >= cutoff)

ggplot(aps_trend_plot, aes(x = date, y = n, colour = country)) +
  geom_point(size = 2, alpha = 0.8) +
  facet_wrap(~ dx_label, ncol = 3, scales = "free_y") +
  scale_colour_country() +
  scale_x_date(date_labels = "%b %y", date_breaks = "3 months") +
  scale_y_continuous(labels = comma) +
  labs(
    x = NULL, y = "Consultations", colour = NULL,
    caption = "Top 5 diagnoses across both countries combined."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Chart 7.4 — Proportion of prenatal clients under 18 over time

`r percent(pct_prenatal_under18, accuracy = 1)` of prenatal clients overall are under 18.

```{r chart-7-4}
#| fig-width: 9
#| fig-height: 4

pren_trend_plot <- prenatal_trend %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff, n_prenatal >= 5) %>%
  mutate(country = label_country(country))

ggplot(pren_trend_plot, aes(x = date, y = pct_under_18, colour = country)) +
  geom_line(linewidth = 0.7, alpha = 0.8) +
  geom_point(aes(size = n_prenatal), alpha = 0.8) +
  scale_colour_country() +
  scale_size_area(max_size = 6, guide = "none") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, NA)) +
  labs(
    x = NULL, y = "% under 18", colour = NULL,
    caption = "Point size = monthly prenatal client count. Months with < 5 prenatal clients excluded."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Chart 7.5 — Mental health caseload over time

```{r chart-7-5}
#| fig-width: 9
#| fig-height: 4

mh_trend_plot <- mh_time %>%
  mutate(date = ym(year_month)) %>%
  filter(date >= cutoff) %>%
  mutate(country = label_country(country))

ggplot(mh_trend_plot, aes(x = date, y = n_mh_consultations, colour = country)) +
  geom_line(linewidth = 0.8, alpha = 0.8) +
  geom_point(size = 1.8, alpha = 0.8) +
  scale_colour_country() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  labs(
    x = NULL, y = "Consultations", colour = NULL,
    caption = "Mental health consultations only. Mexico not included."
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
