---
title: "IRC Latin America — Health Services Dashboard"
---

```{r setup}
library(tidyverse)
library(lubridate)
library(scales)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(tidytext)
library(htmltools)

source("scripts/theme_irc.R")

last_updated_raw <- readLines("data/last_updated.txt")
date_str         <- str_extract(last_updated_raw, "[0-9]{4}-[0-9]{2}-[0-9]{2}")
last_updated_fmt <- paste0(
  "Last updated: ", format(as.Date(date_str), "%d %B %Y"),
  ". Source: CommCare, IRC Latin America."
)

# ── Data loading ──────────────────────────────────────────────────────────────

monthly <- read_csv("data/agg_monthly_consultations.csv", show_col_types = FALSE)
sex_bdown <- read_csv("data/agg_sex_breakdown.csv", show_col_types = FALSE)
pyramid <- read_csv("data/agg_age_sex_pyramid.csv", show_col_types = FALSE)
age_bdown <- read_csv("data/agg_age_breakdown.csv", show_col_types = FALSE)
migration <- read_csv("data/agg_migration_profile.csv", show_col_types = FALSE)
ethnicity <- read_csv("data/agg_ethnic_profile.csv", show_col_types = FALSE)
nationality <- read_csv("data/agg_nationality.csv", show_col_types = FALSE)
svc_dist <- read_csv("data/agg_services_per_client.csv", show_col_types = FALSE)
svc_summary <- read_csv("data/agg_services_per_client_summary.csv", show_col_types = FALSE)
loc_counts <- read_csv("data/agg_location_consultations.csv", show_col_types = FALSE)
loc_coords <- read_csv("data/locations.csv", show_col_types = FALSE)
aps_dx <- read_csv("data/agg_aps_diagnoses.csv", show_col_types = FALSE)
aps_age <- read_csv("data/agg_aps_by_age.csv", show_col_types = FALSE)
chronic <- read_csv("data/agg_chronic_conditions.csv", show_col_types = FALSE)
equity_dx <- read_csv("data/agg_equity_diagnoses.csv", show_col_types = FALSE)
ssr_mix <- read_csv("data/agg_ssr_service_mix.csv", show_col_types = FALSE)
its_results <- read_csv("data/agg_its_results.csv", show_col_types = FALSE)
prenatal_age <- read_csv("data/agg_prenatal_age.csv", show_col_types = FALSE)
cacu <- read_csv("data/agg_cacu_screening.csv", show_col_types = FALSE)
mh_service <- read_csv("data/agg_mh_service_type.csv", show_col_types = FALSE)
mh_dx <- read_csv("data/agg_mh_diagnoses.csv", show_col_types = FALSE)
referrals <- read_csv("data/agg_referrals.csv", show_col_types = FALSE)

# ── Helpers ───────────────────────────────────────────────────────────────────

label_country <- function(x) str_to_title(x)
PROJECT_START <- as.Date("2024-06-01")

# ICD-10 translations
dx_lookup <- c(
  B829 = "Intestinal parasitosis",    N390 = "Urinary tract infection",
  J00  = "Common cold",               Z001 = "Child health check",
  R51  = "Headache",                  J039 = "Acute tonsillitis",
  Z718 = "Other specified consultation", J22 = "Acute lower resp. infection",
  B86  = "Scabies",                   M545 = "Low back pain",
  Z308 = "Contraception care",        J029 = "Acute pharyngitis",
  K029 = "Dental caries",             K297 = "Gastritis",
  J40  = "Bronchitis",                D509 = "Iron deficiency anaemia",
  J459 = "Asthma",                    F99  = "Mental disorder, unspecified",
  Z659 = "Psychosocial problems",     Z639 = "Primary support group problem",
  F419 = "Anxiety disorder, unspecified", F320 = "Mild depressive episode",
  F439 = "Acute stress reaction",     F329 = "Depressive episode, unspecified",
  F321 = "Moderate depressive episode", Z609 = "Social environment problem",
  F322 = "Severe depressive episode", F431 = "PTSD",
  Z619 = "Adverse experience problem", R448 = "Perceptual disturbance",
  Z729 = "Lifestyle problem",         Z625 = "Child neglect problem",
  R453 = "Demoralisation / apathy",   Z733 = "Stress-related problem"
)

label_dx <- function(code, label, maxlen = 30) {
  translated <- dx_lookup[code]
  coalesce(translated, str_trunc(str_to_sentence(label), maxlen))
}

condition_labels <- c(sifilis = "Syphilis", vih = "HIV", hepatitis = "Hepatitis B")

migration_labels <- c(
  poblacion_acogida                     = "Host community",
  migrante_en_transito                  = "Transit migrant",
  migrante_asentado                     = "Settled migrant",
  "migrante_con_vocacin_de_permanencia" = "Migrant (permanent intent)",
  local_transito                        = "Local (in transit)",
  refugiado                             = "Refugee / asylum seeker",
  desplazado_interno                    = "Internally displaced",
  retornado                             = "Returnee / deportee"
)

nationality_labels <- c(
  ecuatoriana        = "Ecuadorian",
  venezolana         = "Venezuelan",
  peruana            = "Peruvian",
  colombiana         = "Colombian",
  otra_nacionalidad  = "Other",
  doble_nacionalidad = "Dual"
)

pop_labels <- c(local_host = "Host community", migrant_refugee = "Migrants / refugees")

chronic_labels <- c(
  hipertension = "Hypertension",
  diabetes     = "Diabetes",
  otro_cronico = "Other chronic",
  cancer       = "Cancer",
  cardiopatias = "Cardiac disease",
  acv          = "Stroke (ACV)",
  epoc         = "COPD"
)

# ── Headline stats ────────────────────────────────────────────────────────────

n_clients_total   <- loc_counts %>% summarise(n = sum(n_clients)) %>% pull(n)
n_clients_ecu     <- loc_counts %>% filter(country == "ecuador") %>% summarise(n = sum(n_clients)) %>% pull(n)
n_clients_per     <- loc_counts %>% filter(country == "peru")    %>% summarise(n = sum(n_clients)) %>% pull(n)

n_services_total  <- monthly %>% pull(n_consultations) %>% sum()
n_services_ecu    <- monthly %>% filter(country == "ecuador") %>% pull(n_consultations) %>% sum()
n_services_per    <- monthly %>% filter(country == "peru")    %>% pull(n_consultations) %>% sum()

pct_under18 <- age_bdown %>%
  filter(age_band %in% c("0-5","6-11","12-17")) %>%
  summarise(pct = sum(n) / sum(age_bdown$n)) %>% pull(pct)

pct_prenatal_under18 <- prenatal_age %>%
  mutate(is_u18 = age_band %in% c("0-5","6-11","12-17")) %>%
  summarise(pct = sum(n[is_u18]) / sum(n)) %>% pull(pct)

syphilis_positivity <- its_results %>%
  filter(condition == "sifilis") %>%
  summarise(rate = sum(n_positive) / sum(n_tested)) %>% pull(rate)

mean_svc_ecu <- svc_summary %>% filter(country == "ecuador") %>% pull(mean_services)
mean_svc_per <- svc_summary %>% filter(country == "peru")    %>% pull(mean_services)
pct_one_ecu  <- svc_summary %>% filter(country == "ecuador") %>% pull(pct_one_service)
pct_one_per  <- svc_summary %>% filter(country == "peru")    %>% pull(pct_one_service)

# ── Value box helper ──────────────────────────────────────────────────────────
# Renders Bootstrap stat cards. `accent` is the top-border colour.

stat_box <- function(value, label, sub = NULL, accent = "#444444") {
  div(
    class = "col",
    div(
      class = "card h-100 text-center p-3",
      style = paste0("border-top: 4px solid ", accent, "; border-radius: 6px;"),
      div(class = "fw-bold mb-1",
          style = paste0("font-size: 1.8rem; color: ", accent, ";"),
          value),
      div(class = "text-muted", style = "font-size: 0.85rem; line-height: 1.3;", label),
      if (!is.null(sub)) div(class = "text-muted mt-1",
                             style = "font-size: 0.75rem;", sub)
    )
  )
}

stat_row <- function(...) div(class = "row row-cols-2 row-cols-md-4 g-3 mb-4", ...)
```

`r last_updated_fmt`

---

## Geographic overview

Where IRC health services are being provided. Bubble size is proportional to total registered clients at each location.

```{r chart-0-1}
#| fig-width: 9
#| fig-height: 6.5

map_data <- loc_counts %>%
  left_join(select(loc_coords, municipio, lat, lon), by = "municipio") %>%
  filter(!is.na(lat), !is.na(lon)) %>%
  mutate(country_label = label_country(country))

region_sf <- ne_countries(
  continent = "South America", scale = "medium", returnclass = "sf"
)
locs_sf <- st_as_sf(map_data, coords = c("lon", "lat"), crs = 4326)

ggplot() +
  geom_sf(data = region_sf, fill = "grey94", colour = "grey72", linewidth = 0.2) +
  geom_sf(data = locs_sf,
          aes(size = n_clients, colour = country_label),
          alpha = 0.75) +
  scale_size_area(max_size = 18, guide = "none") +
  scale_colour_country(name = NULL) +
  coord_sf(xlim = c(-90, -60), ylim = c(-15, 5.0), expand = FALSE, datum = NA) +
  labs(caption = "Bubble size = total registered clients. Municipality centroids.") +
  theme_irc() +
  theme(
    axis.text  = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    legend.position = "bottom"
  )
```

---

## Clients

```{r clients-value-boxes}
#| output: asis
stat_row(
  stat_box(comma(n_clients_total), "Total registered clients",
           accent = "#444444"),
  stat_box(comma(n_clients_ecu),   "Registered in Ecuador",
           sub = "84% Ecuadorian nationals", accent = pal_country["Ecuador"]),
  stat_box(comma(n_clients_per),   "Registered in Peru",
           sub = "55% Venezuelan nationals", accent = pal_country["Peru"]),
  stat_box(percent(pct_under18, accuracy = 1), "Under 18",
           sub = "across both countries", accent = "#E53935")
)
```

### Consultations over time

```{r chart-1-1}
#| fig-width: 9
#| fig-height: 4

monthly_clean <- monthly %>%
  mutate(date = ym(year_month),
         country = label_country(country)) %>%
  filter(date >= PROJECT_START)

ggplot(monthly_clean, aes(x = date, y = n_consultations, colour = country)) +
  geom_line(linewidth = 0.8, alpha = 0.85) +
  geom_point(size = 1.8, alpha = 0.85) +
  scale_colour_country() +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = comma, limits = c(0, NA)) +
  labs(x = NULL, y = "Consultations", colour = NULL) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Sex breakdown

```{r chart-1-2}
#| fig-width: 8
#| fig-height: 3

sex_labels <- c(H = "Men", M = "Women", I = "Intersex", NR = "Prefer not to say", Otro = "Other")

sex_plot <- sex_bdown %>%
  mutate(
    sex_label = sex_labels[sex],
    sex_label = factor(sex_label, levels = c("Men", "Women", "Other", "Intersex", "Prefer not to say")),
    country   = label_country(country),
    pct_label = str_c(round(pct * 100, 1), "%")
  ) %>%
  filter(sex_label != "Other")

ggplot(sex_plot, aes(x = sex_label, y = 1, colour = sex_label)) +
#  geom_col(alpha = 0.8) +
  geom_point(aes(size = pct)) +
#  geom_text(aes(label = pct_label), colour = "black", size = 3) +
  geom_text(data = filter(sex_plot, sex_label == "Women"), 
            aes(label = pct_label), colour = "white", size = 3.5) +
  geom_text(data = filter(sex_plot, sex_label == "Men"), 
            aes(label = pct_label), colour = "black", size = 3.5) +
#  facet_wrap(~ country, ncol = 2) +
  facet_grid(cols = vars(country)) +
#  scale_x_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_colour_sex() +
  scale_size_area(max_size = 30, guide = "none") +
  labs(x = NULL, y = NULL) +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    panel.grid.major = element_blank()
  )
```

### Age and sex distribution

`r percent(pct_under18, accuracy = 1)` of all registered clients are under 18.

```{r chart-1-3}
#| fig-width: 9
#| fig-height: 5.5

age_bands_5yr  <- c("0-4","5-9","10-14","15-19","20-24","25-29",
                    "30-34","35-39","40-44","45-49","50-54","55-59","60-64","65+")
age_labels_5yr <- str_replace(age_bands_5yr, "-", "–")
names(age_labels_5yr) <- age_bands_5yr

pyramid_plot <- pyramid %>%
  mutate(
    sex_label = if_else(sexo == "M", "Women", "Men"),
    value     = if_else(sexo == "M", pct, -pct),
    age_5yr   = factor(age_5yr, levels = age_bands_5yr),
    country   = label_country(country)
  )
x_limit <- max(abs(pyramid_plot$value)) * 1.08

ggplot(pyramid_plot, aes(x = value, y = age_5yr, fill = sex_label)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2) +
  scale_x_continuous(limits = c(-x_limit, x_limit),
                     labels = function(x) percent(abs(x), accuracy = 0.1)) +
  scale_y_discrete(labels = age_labels_5yr) +
  scale_fill_sex() +
  labs(x = NULL, y = NULL, fill = NULL,
       caption = "% of total clients by sex. Intersex and prefer-not-to-say excluded (< 0.1% combined).") +
  theme(legend.position = "top")
```

### Migration profile

Ecuador's clients are predominantly host community (77%); Peru's are predominantly transit migrants (69%) — reflecting fundamentally different programme contexts.

```{r chart-1-4}
#| fig-width: 9
#| fig-height: 4.5

migration_plot <- migration %>%
  mutate(
    profile_label = coalesce(migration_labels[migration_profile], migration_profile),
    country       = label_country(country)
  ) %>%
  group_by(country) %>%
  mutate(profile_label = reorder(profile_label, pct)) %>%
  ungroup()

ggplot(migration_plot, aes(x = pct, y = profile_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2, scales = "free_y") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL)
```

### Ethnic profile

```{r chart-1-5}
#| fig-width: 8
#| fig-height: 4

pct_indigena <- ethnicity %>%
  filter(ethnic_profile == "indigena") %>%
  summarise(pct = sum(n) / sum(ethnicity$n)) %>% pull(pct)

ethnic_labels <- c(
  mestizo                 = "Mestizo",
  indigena                = "Indigenous",
  afrodescendiente        = "Afro-descendant",
  caucsico                = "White",
  sin_autorreconocimiento = "Other",
  montubio                = "Montubio",
  otro_ee                 = "Other",
  otro_ep                 = "Other"
)

ethnic_plot <- ethnicity %>%
  mutate(
    ethnic_label = coalesce(ethnic_labels[ethnic_profile], ethnic_profile),
    ethnic_label = case_when(
      ethnic_label %in% c("Other", "No self-identification", "Unknown") ~ "Other",
      TRUE ~ ethnic_label
    ),
    country = label_country(country)
  ) %>%
  group_by(country, ethnic_label) %>%
  summarise(n = sum(n), pct = sum(pct), .groups = "drop") %>%
  group_by(country) %>%
  mutate(
#    ethnic_label = reorder(ethnic_label, pct),
    pct_label = str_c(round(pct * 100, 1), "%")
  ) %>%
  ungroup()

ggplot(ethnic_plot, aes(x = pct, y = ethnic_label, fill = country)) +
  geom_col(alpha = .8) +
  geom_text(
    aes(label = pct_label),
    hjust = -0.3,
    colour = "black", size = 3
  ) +
  facet_grid(cols = vars(country)) +
  scale_fill_country(guide = "none") +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.2))
  ) +
  labs(x = NULL, y = NULL)
```

### Nationality

Ecuador is predominantly serving its own nationals (84%) alongside Venezuelan and Colombian migrants. Peru's caseload is majority Venezuelan (55%), reflecting its role as a transit corridor.

```{r chart-1-6}
#| fig-width: 8
#| fig-height: 4

nat_plot <- nationality %>%
  mutate(
    nat_label = coalesce(nationality_labels[nacionalidad], nacionalidad),
    country   = label_country(country),
    pct_label = str_c(round(pct * 100, 1), "%")
  )

ggplot(nat_plot, aes(x = pct, y = nat_label, fill = country)) +
  geom_col(alpha = .8) +
  geom_text(
    aes(label = pct_label),
    hjust = -0.3,
    colour = "black", size = 3
  ) +
  facet_grid(cols = vars(country)) +
  scale_fill_country(guide = "none") +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0.3))
  ) +
  labs(x = NULL, y = NULL)
```

### Services per client

`r percent(pct_one_ecu, accuracy = 1)` of Ecuador clients and `r percent(pct_one_per, accuracy = 1)` of Peru clients received only a single service. Peru's higher return rate (mean `r mean_svc_per` vs `r mean_svc_ecu` in Ecuador) is consistent with a more complex caseload in a transit setting.

```{r chart-1-7}
#| fig-width: 8
#| fig-height: 4

band_order <- c("1", "2", "3", "4-5", "6-10", "11+")

svc_plot <- svc_dist %>%
  mutate(
    n_services_band = factor(n_services_band, levels = band_order),
    country         = label_country(country)
  )

ggplot(svc_plot, aes(x = n_services_band, y = pct, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_country(guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Services received", y = NULL,
       caption = "Total services per registered client, all time.")
```

---

## Services

```{r services-value-boxes}
#| output: asis
stat_row(
  stat_box(comma(n_services_total), "Total services delivered",
           accent = "#444444"),
  stat_box(comma(n_services_ecu),   "Services in Ecuador",
           accent = pal_country["Ecuador"]),
  stat_box(comma(n_services_per),   "Services in Peru",
           accent = pal_country["Peru"]),
  stat_box(percent(syphilis_positivity, accuracy = 0.1), "Syphilis positivity",
           sub = "Peru 5.9% · Ecuador 1.8%", accent = "#E53935")
)
```

---

### Section 2: Primary care

#### Top 10 diagnoses by country

Ecuador's top diagnoses reflect a settled community caseload (parasitosis, UTI, cold). Peru's prominence of child health checks (Z00.1) and headache suggests a different service model — more health surveillance in a transit population with limited prior care.

```{r chart-2-1}
#| fig-width: 10
#| fig-height: 5.5

dx_top10 <- aps_dx %>%
  mutate(dx_label = label_dx(diagnosis_code, diagnosis_label)) %>%
  group_by(country) %>%
  slice_max(n, n = 10) %>%
  mutate(
    dx_label = reorder(dx_label, n),
    country  = label_country(country)
  ) %>%
  ungroup()

ggplot(dx_top10, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = comma) +
  labs(x = "Consultations", y = NULL,
       caption = "Top 10 diagnoses per country by total consultation count.")
```

#### Diagnoses by age band

```{r chart-2-2}
#| fig-width: 11
#| fig-height: 13

age_order <- c("0-5","6-11","12-17","18-49","50+")

dx_by_age <- aps_age %>%
  mutate(
    dx_label = label_dx(diagnosis_code, diagnosis_label),
    age_band = factor(age_band, levels = age_order),
    country  = label_country(country)
  ) %>%
  group_by(country, age_band) %>%
  slice_max(n, n = 5) %>%
  mutate(dx_label = reorder_within(dx_label, n, list(country, age_band))) %>%
  ungroup()

ggplot(dx_by_age, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_grid(age_band ~ country, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  labs(x = "Consultations", y = NULL,
       caption = "Top 5 diagnoses per country and age band.") +
  theme(axis.text.y  = element_text(size = 8),
        strip.text.y = element_text(size = 9))
```

#### Chronic conditions

Among the 44% of clients with chronic conditions data, hypertension and diabetes account for the majority of named conditions. The high share of `otro_cronico` reflects free-text entries that were redacted at anonymisation.

```{r chart-2-3}
#| fig-width: 8
#| fig-height: 4

chronic_plot <- chronic %>%
  mutate(
    cond_label = coalesce(chronic_labels[condition], condition),
    cond_label = reorder(cond_label, n),
    country    = label_country(country)
  )

ggplot(chronic_plot, aes(x = pct, y = cond_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2, scales = "free_y") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 0.1)) +
  labs(x = NULL, y = NULL,
       caption = "% of clients with chronic conditions data (44% of total). Clients may have more than one condition.")
```

#### Diagnoses: host community vs migrants

```{r chart-2-4}
#| fig-width: 10
#| fig-height: 7

eq_plot <- equity_dx %>%
  mutate(
    dx_label  = label_dx(diagnosis_code, diagnosis_label),
    pop_label = coalesce(pop_labels[population_group], population_group),
    country   = label_country(country)
  ) %>%
  group_by(country, pop_label) %>%
  mutate(dx_label = reorder_within(dx_label, n, list(country, pop_label))) %>%
  ungroup()

ggplot(eq_plot, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_grid(pop_label ~ country, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_y_reordered() +
  scale_x_continuous(labels = comma) +
  labs(x = "Consultations", y = NULL,
       caption = "Top 5 diagnoses per country and population group.") +
  theme(axis.text.y = element_text(size = 8))
```

---

### Section 3: Sexual and reproductive health

#### SSR service mix

```{r chart-3-1}
#| fig-width: 7
#| fig-height: 4

ssr_labels <- c(
  planificacin_familiar = "Family planning",
  atencin_prenatal      = "Prenatal care",
  its                   = "ITS/STI",
  despistaje_de_cancer  = "Cancer screening",
  vbg                   = "GBV services",
  consejera_en_ssr      = "SSR counselling"
)

ssr_plot <- ssr_mix %>%
  mutate(
    service_label = coalesce(ssr_labels[ssr_tipo], ssr_tipo),
    country       = label_country(country)
  )

ggplot(ssr_plot, aes(x = country, y = pct, fill = service_label)) +
  geom_col(alpha = 0.85) +
  scale_fill_manual(values = pal_service, name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL,
       caption = "Proportion of SSR consultations by service type.") +
  theme(legend.position = "right")
```

#### ITS/STI testing and positivity

Syphilis positivity is **`r percent(syphilis_positivity, accuracy = 0.1)`** combined. Peru (5.9%) is substantially higher than Ecuador (1.8%) — a clinically significant difference with implications for testing protocols and treatment supply.

```{r chart-3-2}
#| fig-width: 8
#| fig-height: 4

its_plot <- its_results %>%
  mutate(
    condition_label = coalesce(condition_labels[condition], condition),
    condition_label = factor(condition_label, levels = c("Syphilis","HIV","Hepatitis B")),
    country         = label_country(country)
  )

ggplot(its_plot, aes(x = condition_label, y = positivity_rate, fill = country)) +
  geom_col(alpha = 0.85, position = position_dodge(0.7), width = 0.6) +
  geom_text(
    aes(label = paste0("n=", n_tested), y = positivity_rate + 0.003),
    position = position_dodge(0.7),
    size = 3, vjust = 0, family = "Inter"
  ) +
  scale_fill_country(name = NULL) +
  scale_y_continuous(labels = percent_format(accuracy = 0.1), limits = c(0, NA)) +
  labs(x = NULL, y = "Positivity rate",
       caption = "n = number tested. Hepatitis B sample sizes are small — interpret with caution.")
```

#### Family planning method mix

*Data not available.* The `fp_corta` (short-acting) and `fp_larga` (long-acting) fields are empty across the full EPR dataset. Family planning consultations are counted in the SSR service mix above; method breakdown pending field-name investigation.

#### Prenatal care: age profile

`r percent(pct_prenatal_under18, accuracy = 1)` of prenatal clients are under 18.

```{r chart-3-4}
#| fig-width: 7
#| fig-height: 4

age_order    <- c("0-5","6-11","12-17","18-49","50+")
age_display  <- c("0–5","6–11","12–17","18–49","50+")
names(age_display) <- age_order

prenatal_plot <- prenatal_age %>%
  mutate(
    age_band = factor(age_band, levels = age_order),
    bar_fill = if_else(age_band %in% c("0-5","6-11","12-17"),
                       col_highlight,
                       pal_country[str_to_title(country)]),
    country  = label_country(country)
  )

ggplot(prenatal_plot, aes(x = age_band, y = pct, fill = bar_fill)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_identity() +
  scale_x_discrete(labels = age_display) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL,
       caption = "Red bars = under 18. Age at registration.")
```

#### Cervical cancer screening (CACU/IVAA)

```{r chart-3-5}
#| output: asis

cacu_ecu <- cacu %>% filter(country == "ecuador")

cat(paste0(
  "**Ecuador:** ", cacu_ecu$n_screened_ivaa, " clients screened by IVAA. ",
  "Positivity rate: ", percent(cacu_ecu$ivaa_positivity_rate, accuracy = 0.1),
  " (", cacu_ecu$n_ivaa_positive, " positive).",
  "\n\n",
  "**Peru:** No IVAA screening records in current dataset. ",
  "Confirm whether the service runs in Peru or whether the field name differs."
))
```

---

### Section 4: Mental health

*Mental health services are provided in Peru and Ecuador only.*

#### Mental health service type

```{r chart-4-1}
#| fig-width: 8
#| fig-height: 3.5

mh_service_labels <- c(
  atencin_psicsocial_basada_en_la_comunidad                  = "Community psychosocial",
  atencin_psicosocial_focalizada                             = "Focused psychosocial",
  "atencin_especializada_referencia_a_atencin_especializada" = "Specialised / referred"
)

mh_svc_plot <- mh_service %>%
  mutate(
    svc_label = coalesce(mh_service_labels[service_type], service_type),
    svc_label = reorder(svc_label, pct),
    country   = label_country(country)
  )

ggplot(mh_svc_plot, aes(x = pct, y = svc_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2) +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL)
```

#### Top mental health diagnoses

```{r chart-4-2}
#| fig-width: 9
#| fig-height: 5

mh_top <- mh_dx %>%
  mutate(dx_label = label_dx(diagnosis_code, diagnosis_label)) %>%
  group_by(country) %>%
  slice_max(n, n = 10) %>%
  mutate(
    dx_label = reorder(dx_label, n),
    country  = label_country(country)
  ) %>%
  ungroup()

ggplot(mh_top, aes(x = n, y = dx_label, fill = country)) +
  geom_col(alpha = 0.85) +
  facet_wrap(~ country, ncol = 2, scales = "free") +
  scale_fill_country(guide = "none") +
  scale_x_continuous(labels = comma) +
  labs(x = "Consultations", y = NULL)
```

#### External referral rate

```{r chart-4-3}
#| fig-width: 5
#| fig-height: 3

ref_plot <- referrals %>% mutate(country = label_country(country))

ggplot(ref_plot, aes(x = country, y = referral_rate, fill = country)) +
  geom_col(alpha = 0.85, width = 0.5) +
  geom_text(
    aes(label = paste0(percent(referral_rate, accuracy = 0.1),
                       "\n(", n_referred, " of ", comma(n_consultations), ")")),
    vjust = -0.3, size = 3.2, family = "Inter"
  ) +
  scale_fill_country(guide = "none") +
  scale_y_continuous(labels = percent_format(accuracy = 0.1),
                     limits = c(0, max(ref_plot$referral_rate) * 1.5)) +
  labs(x = NULL, y = "Referral rate",
       caption = "Proportion of mental health consultations with external specialist referral.")
```

---

### Section 5: Pharmaceutical dispensing

*Mexico pipeline not yet built. This section will be populated in a future update.*

---

*`r last_updated_fmt` · Ecuador and Peru data only. Mexico deferred.*
